---
title: "Report"
author: "Team _"
date: "2024-01-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load, include = FALSE}
Sys.setlocale("LC_TIME", "C")
library(timeSeries)

# Load Data
load("aeso.RData")

# Load Covariates
weather = readRDS("weather.rds")
```

## NA's

These are all the records with missing values:

```{r NAs, echo=FALSE}
rows_with_missing_values <- aeso[is.na(aeso[['Edmonton']]), ]
print(rows_with_missing_values)
aeso <- na.omit(aeso)
```

This link shows daylight savings dates in Canada, they coincide with the missing values: '<https://www.timetemperature.com/canada/daylight_saving_time_canada.shtml>'

The rest of the data does not have missing values. These data points were removed from the dataset for futher steps as those timestamps do not really exist.

```{r AGG, echo=FALSE}
### Get daily max value
aeso$DT_MST <- as.POSIXct(aeso$DT_MST, format = "%Y-%m-%d %H:%M:%S")
aeso$Date <- as.Date(aeso$DT_MST)
#max <- aggregate(Edmonton ~ Date, data = aeso, FUN = max, na.rm = TRUE)

max <- aggregate(aeso$Edmonton, by = list(Date = aeso$Date), FUN = max, na.rm = TRUE)
colnames(max) <- c("Date", "Edmonton")

### Creating timeSeries
ts = timeSeries(max$Edmonton,
                max$Date, format = "%Y-%m-%d")
```

## Timeseries Plots

```{r Plot1, echo=FALSE, warning=FALSE}
# Whole dataset
plot(ts, ylab="Edmonton hourly demand (in MW)", at="pretty", main = 'Max Daily Demand Edmonton')
```

Observations:

-   Upward trend from 2011 to 2015

-   Outliers to investigate: One point in january 2012

```{r Plot2, echo=FALSE, warning=FALSE}
# Year 2015
plot(window(ts, start=timeDate("2015-01-01", format="%Y-%m-%d"),
            end=timeDate("2015-12-31", format="%Y-%m-%d")),
     ylab="Edmonton Max daily demand 2015",
     xlab="Year 2015", at="pretty", 
     main = 'Max Daily Demand Edmonton 2015')
```

Observations:

-   Seasonal pattern

-   Highest max consumption in summer and winter

-   Highest variablility in summer

```{r Plot3, echo=FALSE, warning=FALSE}
# First week of 2015
plot(window(ts, start=timeDate("2015-01-04", format="%Y-%m-%d"),
            end=timeDate("2015-01-10", format="%Y-%m-%d")),
     ylab="Edmonton max daily demand (in MW)",
     xlab="Sunday to Saturday",
     at="pretty", 
     main = 'First week of 2015')
```

Observations:

-   High max consumption on monday

-   Low ma consumption on saturday

```{r Plot4, echo=FALSE, warning=FALSE}
# Show two different weeks in January 2015
week1 <- series(window(ts,
                       start=timeDate("2015-01-04", format="%Y-%m-%d"),
                       end=timeDate("2015-01-10", format="%Y-%m-%d")))
week2 <- series(window(ts,
                       start=timeDate("2015-01-11", format="%Y-%m-%d"),
                       end=timeDate("2015-01-17", format="%Y-%m-%d")))
plot(week1, axes=F,
     lty=1, type="l", ylim=c(min(week1,week2),max(week1,week2)),
     ylab="Edmonton max daily demand (in MW)", xlab="", 
     main = 'First two weeks of january 2015')
lines(week2, lty=3, lwd=1.8)
axis(1,at =1:7,
     labels=c("Sun","Mon","Tue","Wed","Thu","Fri","Sat"))
axis(2)
legend("bottomleft", legend=c("Jan 4-10, 2015","Jan 11-17, 2015"), 
       lty=c(1,3))

```

Observations:

-   Similar patterns for both weeks

```{r Plot5, echo=FALSE, warning=FALSE}
# Show weekly difference between January and June
week1 <- series(window(ts,
                       start=timeDate("2015-01-04", format="%Y-%m-%d"),
                       end=timeDate("2015-01-10", format="%Y-%m-%d")))
week2 <- series(window(ts,
                       start=timeDate("2015-06-05", format="%Y-%m-%d"),
                       end=timeDate("2015-06-11", format="%Y-%m-%d")))
plot(week1, axes=F,
     lty=1, type="l", ylim=c(min(week1,week2),max(week1,week2)),
     ylab="Edmonton max daily demand (in MW)", xlab="", 
     main = 'January week versus June')
lines(week2, lty=3, lwd=1.8)
axis(1,at =1:7,
     labels=c("Sun","Mon","Tue","Wed","Thu","Fri","Sat"))
axis(2)

legend("bottomright", legend=c("Jan 4-10, 2015","Jun 5-11, 2015"), 
       lty=c(1,3))
```

Observations:

-   Opposite patterns between summer and winter

## Plots covariates

```{r Merge, echo=FALSE, warning=FALSE}
# Merge data
df <- merge(max, weather, by = "Date", all = FALSE)
head(df, n =200)
```

This is an overview of the covariates that could be useful in our analysis. The data was retrieved from Environment Canada API. The data was collected by the Blatchford weather station in Edmonton. Missing values were imputed using nearby weather station (15 km radius). Some variables were engineered like HDD, CDD and other variables derived from them, as well as business day.

-   Max, Min and Mean temperature for a given day retrieved from blatchford station.

-   HDD and CDD (Heating Degree Days and Cooling Degree Days) with 10 degrees as a reference point. The reference point was determined from the plot below. The variables with subscript n (\_n) represent the measures taken at a lag(n). The prefix 'Cons' refers to consecutive heating degree days or cooling degree days.

-   Temporal data like Month, Year, DoW (Day of Week) and Business Day. The DoW indicator was derived from a canadian holiday calendar (not specific to Alberta).

```{r Plot6, echo=FALSE, warning=FALSE}
# Demand versus temperature
plot(df$MeanTemp, df$Edmonton, 
     xlab = "MeanTemp", ylab = "Demand MW",
     main = "Scatterplot of Demand versus Mean Temp")
```

Observations:

-   We clearly see a change of effect of temperature on demand at around 10 degrees.

-   We also observe that there are two different groups of points one over the other. This could be explained by another binary variable.

```{r Plot7, echo=FALSE, warning=FALSE}
# Demand versus HDD (Accounting for ConsHDD)
plot(df$HDD, df$Edmonton, 
     xlab = "HDD", ylab = "Demand MW",
     main = "Scatterplot of Demand versus HDD (ConsHDD)",
     col = ifelse(df$ConsHDD == 1, "red", "blue"))
legend("bottomright", legend = c("ConsHDD = 1 ", "ConsHDD = 0"), 
       col = c("red", "blue"), pch = 1)
```

Observations:

-   Consecutive Heating degree days does not seem to provide much additional information on this graph.

```{r Plot8, echo=FALSE, warning=FALSE}
# Demand versus HDD (Accounting for business day)
plot(df$HDD, df$Edmonton, 
     xlab = "HDD", ylab = "Demand MW",
     main = "Scatterplot of Demand versus HDD (BusinessDay)",
     col = ifelse(df$IsBusinessDay == 1, "red", "blue"))
legend("bottomright", legend = c("IsBusinessDay = 1", "IsBusinessDay = 0"), 
       col = c("red", "blue"), pch = 1)
```

Observations:

-   Business day seems to be a very good variable that segregates between low and high demand for a given temperature.

```{r Plot9, echo=FALSE, warning=FALSE}
# Demand versus CDD (Accounting for ConsCDD)
plot(df$CDD, df$Edmonton, 
     xlab = "CDD", ylab = "Demand MW",
     main = "Scatterplot of Demand versus CDD (ConsCDD)",
     col = ifelse(df$ConsCDD == 1, "blue", "red"))
legend("bottomright", legend = c("ConsCDD = 1 ", "ConsCDD = 0"), 
       col = c("blue","red"), pch = 1)
```

Observations:

-   Here we clearly see the two curves one over the other.

```{r Plot10, echo=FALSE, warning=FALSE}
# Demand versus CDD (Accounting for business day)
plot(df$CDD, df$Edmonton, 
     xlab = "CDD", ylab = "Demand MW",
     main = "Scatterplot of Demand versus CDD (BusinessDay)",
     col = ifelse(df$IsBusinessDay == 1,"red","blue"))
legend("bottomright", legend = c("IsBusinessDay = 1", "IsBusinessDay = 0"), 
       col = c("red","blue"), pch = 1)
```

Observations:

-   Business day explains very well the differences between the curves.


## TO DO
• an introduction with a good description of the zone under study
• exploratory data analysis
• evaluation of naive methods
• description of possible explanatory variables, clearly indicating the source of the data,
and providing some summary statistics

## Format
Part I of the report, in .pdf format:
– with AT MOST 12 PAGES. Includes all submitted pages except cover page2
;
– in Times New Roman font, 12 pt, double-spaced, “letter” page format with minimum 1 inch margins.

